\section{Related Work}
\label{sec:related}

\paragraph{Prior coordination systems.}
CA-MCP~\cite{camcp2026} is the nearest prior system by name and
substrate: server-to-server, positive plan state, single-trial;
the authors explicitly disclaim cross-trial use.
Terrarium~\cite{terrarium2025} is a second MCP-based system
with a pull-style interface, but targets attack-surface analysis.
MPAC~\cite{mpac2026} proposes Lamport-clock conflict resolution
for the multi-principal case. Reflexion and
ExpeL~\cite{reflexion2023,expel2024} share the negative-knowledge
intuition but inject via prompt context and are single-agent or
cross-task-transfer. Long-lived cross-task memory
(MemGPT~\cite{memgpt2024}, G-Memory~\cite{gmemory2025}) operates
at task-spanning rather than within-task granularity.
AgentVerse and related orchestration substrates surveyed
in~\cite{mast2025} round out the prior-systems comparison
revisited in Table~\ref{tab:literature-vs-floor}. Framework-level hooks
(Strands~\cite{strandshooks2025}, AgentCore
Policy~\cite{agentcorepolicy2025},
LangChain middleware~\cite{langchainmiddleware2026},
Microsoft Agent Framework~\cite{msagentframework2026},
AutoGen~\cite{autogen2023}, MetaGPT~\cite{metagpt2024})
carry blocking, validation, retries, and caching but no
coordination semantics, and any of them could host the same
measurement protocol.

\paragraph{Measurement methodology: the actual gap.}
A concurrent thread measures the reliability of agent
benchmarks. Yu~et~al.~\cite{multiagentmem2026} frame the
multi-agent case as lacking remote direct memory access.
Variance-decomposition work spans a 12-metric
framework~\cite{rabanser2026reliability}, an ICC
framework~\cite{mustahsan2025stochasticity}, Anthropic-side
infrastructure noise~\cite{anthropicnoise2026}, aggregate
pass@1 variance~\cite{randomnessagentic2026}, systematic
agent reliability~\cite{reliabilitybench2026}, and input-noise
robustness~\cite{agentnoisebench2026}.
\textbf{None of these reports paired configuration-equivalent
replication on a state-validated multi-agent coordination
benchmark, and none conditions on a coordination-active
sub-event.} Bedi~et~al.~\cite{bedi2025paradox} (KDD~'25,
same workshop series) document a component-vs-system paradox
in clinical multi-agent settings but treat the system as a
black box. Cuadron~et~al.'s
$\tau^2$-bench-verified~\cite{tau2bench_verified2025} fixes
tasks while ours fixes transport; the two compose. AgentTrace's
post-hoc causal graphs~\cite{agenttrace2026} and
STATE-Bench~\cite{statebench2026}'s memory-agnostic harness
are complementary rather than competitive.
The central claim of this paper, that coordination effect
sizes should be reported relative to a same-model paired
coordination-active variance gate, is what fills that gap.
Ou~et~al.~\cite{ou2025coordination} compare
orchestrator-mediated vs.\ shared-notebook coordination on
TravelPlanner but neither pair at matched payload nor report
a coordination-active variance floor, and we revisit them in
Table~\ref{tab:literature-vs-floor} as one of the recent
coordination papers whose reported effects straddle the paired
noise envelope measured here on Haiku~4.5 / $\tau^2$-bench
retail (pooled upper CI $\lesssim 15$\,pp).
