\section{Forward Work}
\label{sec:proposals}
\label{sec:discussion}

The coordination-active null
(\S\ref{sec:eval:results:retail-intercept}) isolated three failure
modes of naive coordination, each with a matched refinement on
existing production hook surfaces (Strands~\cite{strandshooks2025},
AgentCore Policy~\cite{agentcorepolicy2025}, LangChain
middleware~\cite{langchainmiddleware2026}).
\textbf{P1: causally-attributed writers} address \emph{M1
writer mis-attribution}: an LLM-judge consumes the trajectory
and reward and re-weights which trial calls get written, replacing
the recency-only last-$K$ heuristic.
\textbf{P2: selective intercept} addresses \emph{M2 per-match
injection noise}: a score gate fires only when a peer event is
relevant, novel, and decision-proximal (preliminary signal below).
\textbf{P3: predicate constraint extraction} addresses \emph{M3
brittle literal keying}: an extractor turns the trajectory into a
typed predicate (e.g.\ \texttt{tool=cancel\_order} with
\texttt{order.status=shipped}) so intercept fires on the
predicate, not the literal argument tuple.
\textbf{P4: coordination-active pass$^k$}
(\S\ref{sec:eval:metrics}) is the release-time gate any refinement
should clear. P1 and P3 each require an LLM-judge component not
validated here. A shared trace store also opens security surfaces
(poisoned writes, cross-principal exfiltration); the prototype
assumes a trusted single-tenant deployment.

\paragraph{P2 (selective intercept), preliminary signal.}
A score-gated intercept that fires only when a peer event is
relevant, novel, and decision-proximal beat vanilla intercept on
the coordination-active subset at Haiku~4.5 retail $n{=}30$
($+15.8$\,pp, paired sign $2/7/21$, $p{=}0.18$; hyperparameters
not pre-registered). Directionally consistent and underpowered;
Sonnet at $n{\geq}250$ is the natural confirmatory test.
