
\section{Conclusion}
\label{sec:conclusion}

The observed trial-0 paired-gap envelope on $\tau^2$-bench
retail spans $[{-}3,{+}18]$\,pp across all three contrasts and
two $n{=}100$ Haiku 4.5 seeds; the clean
configuration-equivalent contrast (no\_coord vs.\ intercept)
gives $+10$ and $0$\,pp, pooling to $+5$\,pp (CI
$[{-}2,{+}12]$), with a pooled upper Wilson CI of
${\sim}15$\,pp on the largest between-protocol contrast; no
single-seed Bonferroni significance survives a second seed. ET-MCP, the task-scoped
trace-store substrate, made the matched-payload comparison
possible; the measurement discipline, not the substrate, is
what we expect to travel. Notably, our own seed-$1$
pull-vs-intercept $p_{\text{corr}}{=}0.012$ did not reproduce
at seed-$2$---a worked example of the protocol catching the
single-seed artifact it is designed to catch.

Seven of ten recent coordination architectures report headline
gains below that envelope and one more sits inside it. The
coordination-active head-to-head here is an underpowered
non-result, not a clean null: the informative pairs after ties
are $n{=}8$--$17$, well short of what a medium-sized effect
would require, so it reframes the question rather than retiring
it. The negative reading applies to \emph{naive} coordination
under a measurement-honest protocol, not to coordination as a
research program. P1--P3 (\S\ref{sec:proposals}) are the
mechanisms we expect to clear the floor, and
coordination-active pass$^k$ is the gate they should be judged
against. Reporting a same-model paired floor alongside any
multi-agent delta is the practice the field can adopt
now, with the protocol here as a starting template.
