% Shorter, first-read-accessible abstract: problem -> measurement ->
% finding -> metric -> substrate. Detailed numbers moved to body.

Multi-agent LLM coordination papers report small
benchmark deltas as evidence that one architecture beats
another. A prior question: how much paired
trial-$0$ disagreement do two protocols produce on the same
model and benchmark when their API inputs are
\emph{configuration-equivalent} (matched by code inspection
plus a SHA-$256$ byte audit), short of full
identity-replay? On Claude Haiku $4.5$ against
$\tau^2$-bench retail, the clean configuration-equivalent
contrast (no\_coord vs.\ intercept, both inert at trial $0$)
gives signed paired gaps of $+10$\,pp and $0$\,pp across two
$n{=}100$ seeds; pooled across both, $+5$\,pp with Wilson CI
$[{-}2,{+}12]$, not significant. The largest single-seed
contrast ($+18$\,pp pull-vs-intercept, $p_{\text{corr}}{=}0.012$)
did not reproduce at the second seed ($-3$\,pp,
$p_{\text{corr}}{=}1.0$); no trial-$0$ contrast is
significant after Bonferroni at either seed or pooled. The envelope of
observed paired gaps spans $[{-}3,{+}18]$\,pp across two seeds,
with pooled upper Wilson CI $\lesssim 15$\,pp.
\textbf{Seven of ten recent multi-agent coordination
architectures report headline effects below this local floor,
and one more sits inside the envelope}; whether they survive a same-model paired
replication is, by construction, untested in their original
settings.

We define \emph{coordination-active pass$^k$}, pass$^k$
restricted to trials where the coordination mechanism is
logically active, as the minimum reporting protocol, with
sample-size targets and runtime hooks in the body. Measurements run on \textbf{ET-MCP}, a
task-scoped negative-knowledge store conformant with MCP
2026-07-28, used as a substrate to isolate reader-side
choices, not as a contribution. On Haiku $4.5$ the
candidate readers (pull, intercept) do not improve
trial-$1$ recovery; we give a preliminary diagnosis of
failure modes with refinements on existing production hook
surfaces~\cite{strandshooks2025,agentcorepolicy2025,langchainmiddleware2026,sagemaker_monitor2022}.
