\section{Introduction}
\label{sec:introduction}

Two LLM agents work in parallel on the same multi-step itinerary.
The first attempts to book a flight, discovers no inventory,
abandons the path, and moves on. The second, started moments
later by the same orchestrator, attempts the same booking
through the same API and re-discovers the same dead end. The
first agent's failure lived only in its private scratchpad. This
is a structural property of how multi-agent LLM systems (MAS)
currently coordinate~\cite{mast2025,codedelegator2026}. Recent
architectures address it with coordination channels (structured
handoffs, shared memory, server-side state) and report small
benchmark gains as evidence one design beats another.

\paragraph{Why we wrote this paper.}
We initially set out to evaluate two such designs (an
agent-side \emph{pull} channel and a framework-side
\emph{intercept} channel) against a no-coordination baseline on
$\tau^2$-bench retail. The head-to-head shows no detectable
effect at the available power: the coordination-active subsets
contain $n{=}8$--$17$ informative pairs after ties, far short of
what a medium-sized improvement would require. Unpacking that
non-result forced a question we should have asked at the start:
\emph{what is the run-to-run variance between protocols whose
API inputs are equivalent at trial $0$?} A single number is
enough to invert the read of much of the recent coordination
literature. The rest of the paper answers that question and
traces what it implies for our own and others' architectural
claims.

\paragraph{ET-MCP as substrate.}
Measuring variance under matched payload required a
representative coordination substrate, so we built one.
\textbf{ET-MCP} is an Ephemeral Trace store of typed
negative-knowledge events exposed via the Model Context Protocol
(MCP) 2026-07-28~\cite{mcpspec2026}: task-scoped, append-mostly,
with a TF-IDF (term-frequency / inverse-document-frequency)
ranker over $\textsc{failed\_path}$-typed events. Two reader-side
variants ride on the same store. \emph{Pull} makes the agent the
active reader: it calls \texttt{trace.query} when peer state
might inform the next decision. \emph{Intercept} makes the
framework the active reader: it consults the store transparently
on every tool call and prepends a \texttt{[PEER-WARNING]} block
when (name, arguments) match a peer event. Intercept rides on
the same hook surface already deployed for blocking, validation,
retries, and caching in
Strands~\cite{strandshooks2025}, AgentCore
Policy~\cite{agentcorepolicy2025},
LangChain~\cite{langchainmiddleware2026}, and the
Microsoft Agent Framework~\cite{msagentframework2026}.

\paragraph{The result that reframed the paper.}
Running this substrate against its no-coordination baseline
yielded no detectable effect at the available power. At
$\tau^2$-bench retail, Haiku $4.5$, $T{=}0$, $n{=}100$ tasks,
pull, intercept, and no\_coord all reach trial-$1$ success
$0.54$ (paired sign tests fail to reject, $p{=}1.0$ for all
three contrasts;
\S\ref{sec:eval:results:retail-intercept}).
Decomposed by trial, no\_coord and intercept are
\emph{configuration-equivalent} at trial $0$: their requests to
the API are equivalent by code inspection (system prompt, tool
list, sampling parameters, message-array construction; see
\S\ref{sec:eval:request-equivalence} for what we did and did not
verify). This is the only genuinely clean contrast in the
trio; the paired sign-test disagreement is $21/11/68$ at
seed-$1$ and $13/13/74$ at seed-$2$ ($+10$\,pp and $0$\,pp
signed gaps; pooled $+5$\,pp, Wilson CI $[{-}2,{+}12]$,
Bonferroni $p_{\text{corr}}{=}0.711$, not significant). The
largest single-seed paired contrast was pull-vs-intercept at
seed-$1$ ($27/9/64$, $+18$\,pp, $p_{\text{corr}}{=}0.012$),
but it did not reproduce at seed-$2$ ($18/21/61$, $-3$\,pp,
$p_{\text{corr}}{=}1.0$); the pooled gap is
$+7.5$\,pp with Wilson upper CI ${\sim}15$\,pp. We characterize
the configuration-equivalent envelope on this benchmark, model,
and harness as a paired-gap distribution whose observed range
across two seeds is $[{-}3,{+}18]$\,pp and whose pooled upper
Wilson CI is $\lesssim 15$\,pp---no single-seed Bonferroni
significance survives a second seed. The floor is local,
not universal; cross-model and cross-domain probes (Haiku
airline $n{=}30$, Sonnet retail $n{=}30$) confirm it does not
transfer.

\paragraph{Coordination-active pass$^k$.}
To separate run-to-run variance from mechanism effect we define
$\textsc{pass}^k_{\textrm{active}}$: marginal pass$^k$
restricted to tasks where the peer-coordination store was
non-empty at trial start. Marginal pass$^k$ answers ``should I
deploy the whole system?'' Active pass$^k$ answers ``did
coordination help when it had an opportunity to.'' The two
diverge whenever empty-store trials carry unrelated
configuration perturbations, which is the situation we encounter.
Three runtime alarms operationalize the metric on Amazon Bedrock
AgentCore and SageMaker Model
Monitor~\cite{sagemaker_monitor2022}
(\S\ref{sec:eval:deployment}).

\paragraph{Contributions.}
Four contributions track the threads above:
\begin{itemize}
    \item \textbf{(C1) Measurement protocol.} A same-model,
        paired, configuration-equivalent variance procedure for
        coordination-architecture claims on state-validated MAS
        benchmarks. The floor splits into an
        \emph{identity-replay} component (byte-identical
        payloads, which we do not measure) and a
        \emph{configuration-equivalent} component (same intended
        setup, code-inspection-verified, which we
        do)~(\S\ref{sec:eval:request-equivalence}).

    \item \textbf{(C2) Finding.} On $\tau^2$-bench retail with
        Haiku $4.5$ across two $n{=}100$ seeds, the clean
        configuration-equivalent paired-gap pooled CI is
        $[{-}2,{+}12]$\,pp (no\_coord vs.\ intercept);
        no single-seed contrast is significant after Bonferroni
        once a second seed is measured. The observed envelope
        spans $[{-}3,{+}18]$\,pp with pooled upper Wilson CI
        $\lesssim 15$\,pp. Seven of ten recent multi-agent
        coordination architectures
        (Table~\ref{tab:literature-vs-floor})
        report headline gains below this envelope and one more
        sits inside it; each falls below or inside the envelope
        measured here.

    \item \textbf{(C3) Metric.}
        $\textsc{pass}^k_{\textrm{active}}$ (coordination-active
        pass$^k$): pass$^k$ conditioned on a non-empty
        coordination store. Sample-size guidance and three
        runtime instrumentation alarms ship the metric as a
        pre-deployment release gate
        (\S\ref{sec:eval:deployment}).

    \item \textbf{(C4) Failure-mode diagnosis (preliminary).}
        The trial-level data surface three candidate
        failure modes of naive negative-knowledge coordination
        (writer mis-attribution; per-match injection noise;
        brittle keying), with matched architectural refinements
        on existing production hook surfaces
        (\S\ref{sec:proposals}). Single-judge audit and small-$n$
        validation leave these as hypotheses for confirmatory work.
\end{itemize}

ET-MCP itself and the harness mitigating a documented
litellm~\cite{litellmissue12404,litellmissue15322} bug class
appear as supporting artifacts rather than claimed contributions.
