\section{Forward Work}
\label{sec:proposals}
\label{sec:discussion}

The coordination-active null
(\S\ref{sec:eval:results:retail-intercept}) isolated three failure
modes of naive coordination, each with a matched refinement on
existing production hook surfaces (Strands~\cite{strandshooks2025},
AgentCore Policy~\cite{agentcorepolicy2025}, LangChain
middleware~\cite{langchainmiddleware2026}).
\textbf{P1: causally-attributed writers} address \emph{M1
writer mis-attribution}: an LLM-judge consumes the trajectory
and reward and re-weights which trial calls get written, replacing
the recency-only last-$K$ heuristic.
\textbf{P2: selective intercept} addresses \emph{M2 per-match
injection noise}: a score gate fires only when a peer event is
relevant, novel, and decision-proximal (preliminary signal below).
\textbf{P3: predicate constraint extraction} addresses \emph{M3
brittle literal keying}: an extractor turns the trajectory into a
typed predicate (e.g.\ \texttt{tool=cancel\_order} with
\texttt{order.status=shipped}) so intercept fires on the
predicate, not the literal argument tuple.
\textbf{P4: coordination-active pass$^k$}
(\S\ref{sec:eval:metrics}) is the release-time gate any refinement
should clear. P1 and P3 each require an LLM-judge component not
validated here. A shared trace store also opens security surfaces
(poisoned writes, cross-principal exfiltration); the prototype
assumes a trusted single-tenant deployment.

\paragraph{P0: oracle positive control (implemented).}
The protocol shows naive coordination does not clear the floor
but has not yet demonstrated that it \emph{would} detect a real
gain---a measurement-sensitivity gap a reviewer correctly
flagged. We have implemented an \texttt{oracle} arm in the
released harness: it rides pull's exact activation semantics
(inert at trial $0$, so the floor measurement is undisturbed)
but, once a prior trial has failed, injects the task's
\emph{golden action sequence} from the benchmark's evaluation
criteria---perfect coordination content on the same surface.
Coordination-active pass$^k$ against \texttt{no\_coord} on the
oracle arm upper-bounds what any reader-side mechanism could
show on this benchmark. Run at the headline configuration (same
$100$ seed-$42$ tasks, Haiku $4.5$, $T{=}0$), the oracle
confirms trial-$0$ inertness ($-3$\,pp paired gap vs.\
\texttt{no\_coord}, inside the envelope) and lifts recovery on
its trial-$0$-failed subset from $0.35$ to $0.43$
(trial-$1$ marginal $0.60$ vs.\ $0.54$)---but the paired
coordination-active contrast is $8/6/21$ (W/L/T) at $n{=}35$,
$p{=}0.79$: \emph{even ground-truth guidance does not separate
at this $n$}. The direction is right, so the metric reads
real gains; the binding constraint is the subset itself:
$57\%$ of oracle trial-$1$ attempts fail with the golden action
sequence in the system prompt, i.e.\ most coordination-active
failures on this benchmark are not knowledge-limited
(user-simulator dynamics, policy-confirmation flows). This
sharpens the negative reading of
\S\ref{sec:eval:results:retail-intercept}: no reader-side
mechanism could have certified a gain here, because a
perfect-content oracle cannot. Certifying coordination requires
benchmarks whose coordination-active subsets are
knowledge-limited by construction---or the pre-registered
Sonnet $n{\geq}250$ scale.

\paragraph{P2 (selective intercept), preliminary signal.}
A score-gated intercept that fires only when a peer event is
relevant, novel, and decision-proximal beat vanilla intercept on
the coordination-active subset at Haiku~4.5 retail $n{=}30$
($+15.8$\,pp, paired sign $2/7/21$, $p{=}0.18$; hyperparameters
not pre-registered). Directionally consistent and underpowered;
Sonnet at $n{\geq}250$ is the natural confirmatory test.
